---
title: "Visualize Data"
author: "Sequoia Fellenbaum"
format: gfm
execute: 
  echo: true
  eval: true
  warning: false
---


<!-- If you get the error "Error in list2(na.rm = na.rm, ...) : object 'ffi_list2' not found" do the following in the Console run:  

    remove.packages("rlang")

Then Restart R to be safe. Then run:

    install.packages("rlang")

and again restart R to be safe. That should fix the problem (which is a package dependency issue).   -->



Try your code again
```{r}
#| label: setup
#| include: false
library(tidyverse)


```



## Your Turn 0

Add a setup chunk that loads the tidyverse packages.


```{r}
#| label: head-data-mpg
head(mpg)
```


## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
#| label: scatterplot-mpg
ggplot(data = mpg) + 
  geom_point(mapping = aes(x=displ, 
                           y=hwy))

#initializing plot and laying geometry on top
```


## Your Turn 2

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
#| label: boxplot-mpg
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy))
```


## Your Turn 3

Make a histogram of the `hwy` variable from `mpg`. Hint: do not supply a y variable.

```{r}
#| label: histogram-mpg-hwy

ggplot(mpg) +
  geom_histogram(aes(x = hwy))
```

## Your Turn 4

Use the help page for `geom_histogram` to make the bins 2 units wide.

```{r}
#| label: histogram-mpg-hwy-modified

ggplot(mpg) +
  geom_histogram(aes(x = hwy),
                 binwidth = 2,
                 alpha = 0.8, #transparency in color
                 fill = "#245c2f", #fill color
                 color = "black" #outline color
                 )
```


## Your Turn 5

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
#| label: scatterplot-color-by-class-mpg
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, 
                           y = hwy,
                           color = class)) + 
#color was the best option
  labs(x = "Engine size displacement", #x axis title
       y = "Highway mileage", # y axis title
       color = "Class of vehicles", #class color title
       title = "The bigger the engine, the less mileage we get on the highway ") + #graph title
  theme_light() #overall theme
```

## Help Me

What do `facet_grid()` and `facet_wrap()` do? 
(run the code, interpret, convince your group)

```{r}
#| label: exploring-facet-mpg
# Makes a plot that the commands below will modify
q <- ggplot(mpg) + geom_point(aes(x = displ, y = hwy)) #assign base plot to a name for easy iteration testing

q + facet_grid(. ~ cyl) #four columns of graphs - one row, i think organized by cyl
q + facet_grid(drv ~ .) #three rows of graphs - one column, i think organized by drv
q + facet_grid(drv ~ cyl) #twelve graphs - 3 rows by drv 4 by cyl, i think organized by drv horizontally and cyl vertically
q + facet_wrap(~ class) #seven graphs organized by vehicle class
```


## Your Turn 6

Make a bar chart `class` colored by `class`. Use the help page for `geom_bar` to choose a "color" aesthetic for class.

```{r}
#| label: bar-chart-mpg-colored-by-class

ggplot(mpg) +
  geom_bar(aes(x=class,
               fill = class)) + #double-mapping class to two aesthetics
  guides(fill = "none") + #hides legend
  labs(x = "Class of vehicle",
       y = "Number of vehicles in sample",
       title = "Number of vehicles for each class"
  ) +
  theme_bw()
```

```{r}
#| label: bar-chart-mpg-colored-by-drive

ggplot(mpg) +
  geom_bar(aes(x=class,
               fill = drv), 
           position = position_jitter()) + 
  
  labs(x = "Class of vehicle",
       y = "Number of vehicles in sample",
       title = "Number of vehicles for each class",
       fill = "Type of drive"
  ) +
  theme_bw()
```

## Quiz

What will this code do?

Two geometries to one plot

```{r}
#| label: test-code
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy)) 

#ggsave("example.jpg", width = 6, height = 4) #saves most recently rendered plot, will overwrite if changed, commented out to prevent saving every time rendered
```

```{r}
#| label: universal-aesthetics
ggplot(mpg, aes(x=displ, y = hwy)) + 
  geom_point(aes(color = class)) +
  geom_smooth(color = "black",
              se = FALSE, #hides prediction line
              method = "lm") + #linear regression model
  scale_color_viridis_d() #colorblind scale
```


***

# Take aways

You can use this starter code template to make thousands of graphs with **ggplot2**.

```{r}
#| label: template
#| eval: false
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```